\LaTeXe\ guide for the Journal of Functional Programming

### 2019-10-21 papers/jfp/7-experience.tex
Wibbles

Consider, for example, our \hs{Task} type:
\begin{minted}[xleftmargin=10pt]{haskell}
newtype Task c k v = Task (@\std{forall}@ f. c f => (k -> f v) -> f v)
\end{minted}
We started off with a much more concrete type, and explored
multiple variants.  During those iterations, this single line of
code gave us a tangible, concrete basis for productive conversations,
much more so than general conversations about ``tasks''.

It is also worth noting that we needed a rather expressive language,
to faithfully express the abstractions that seem natural in this
setting.  In the case of \hs{Task} we use a data constructor with a
polymorphic field; a higher-kinded type variable \hs{f}~\hs{::}~\hs{*}~\hs{->}~\hs{*}; an
even more abstracted kind for \hs{c}~\hs{::}~\hs{(*}~\hs{->}~\hs{*)}~\hs{->}~\hs{Constraint}; and,
of course, type classes. Our models have been translated to
Rust~\cite{translation_rust} and Kotlin~\cite{translation_kotlin}, and in both
cases there was a loss of precision due to language-specific limitations.

When thinking about the type constructors over which \hs{f} might usefully
range, it turned out that we could adopt the \emph{existing abstractions}
of \hs{Functor}, \hs{Applicative}, \hs{Monad} and so on.
That in turn led us to think in a new way about the taxonomy
of build systems -- see~\S\ref{sec-why-polymorphism}. In the other direction,
trying to express an \emph{existing build system} \Dune in our models led us to
finding a new abstraction -- the \hs{Selective} \mbox{type
class~\cite{mokhov_selective_2019}}, which turned out to be useful outside the
build systems domain.

@@

These powerful and modular abstractions, which ultimately formed part
of the conceptual structure of the paper, emerged fairly late in the
project, as we repeatedly reviewed, redesigned, and refactored our
executable prototypes. It is hard to believe that we could have
developed them without the support of Haskell as a design language.

>>>>

Consider, for example, our \hs{Task} type:
\begin{minted}[xleftmargin=10pt]{haskell}
newtype Task c k v = Task (@\std{forall}@ f. c f => (k -> f v) -> f v)
\end{minted}
We started off with a much more concrete type, and explored
multiple variants.  During those iterations, this single line of
code gave us a tangible, concrete basis for productive conversations,
much more so than general debate about ``tasks''.

It is also worth noting that we needed a rather expressive language
to faithfully express the abstractions that seem natural in this
setting.  In the case of \hs{Task} we needed: a data constructor with a
polymorphic field; a higher-kinded type variable \hs{f}~\hs{::}~\hs{*}~\hs{->}~\hs{*}; an
even more abstracted kind for \hs{c}~\hs{::}~\hs{(*}~\hs{->}~\hs{*)}~\hs{->}~\hs{Constraint}; and,
of course, type classes. Our models have since been translated to
Rust~\cite{translation_rust} and Kotlin~\cite{translation_kotlin}, and in both
cases there was a loss of precision due to language-specific limitations.

When thinking about the type constructors over which \hs{f} might usefully
range, it turned out that we could adopt the \emph{existing} abstractions
of \hs{Functor}, \hs{Applicative}, \hs{Monad} and so on.
That in turn led us to a new taxonomy
of build systems -- see~\S\ref{sec-why-polymorphism}. In the other direction,
trying to express an \emph{existing build system} \Dune in our models led us to
finding a new abstraction -- the \hs{Selective} \mbox{type
class~\cite{mokhov_selective_2019}}, which turned out to be useful outside the
build systems domain.

@@

These powerful and modular abstractions, which ultimately formed part
of the conceptual structure of the paper, emerged fairly late in the
project as we repeatedly reviewed, redesigned, and refactored our
executable prototypes. It is hard to believe that we could have
developed them without the support of Haskell as a design language.
